---
title: "LASSO"
author: "SZ"
date: "`r Sys.Date()`"
output: html_document
---

## Importing libraries and reading data
```{r setup, include=TRUE}
library(glmnet)
library(broom)
library(caret)
combineddata = read.csv("combineddata.csv")

X = model.matrix(sg20y ~ .-Month, data = combineddata)
X=X[, colnames(X)!= "(Intercept)"] 
Y = combineddata$sg20y
```

## Train Test Split
```{r 1, include=TRUE}
train = combineddata[2:143, -1]
train_x = train[,-1]
train_y = train[,1]

test = combineddata[144:177, -1]  # the test sample indices
test_x = test[,-1]
test_y = test[,1]

```


## Setting seed and run model on training data
```{r 2, echo=TRUE,include=TRUE}
set.seed(5)
lasso_mod <- glmnet(train_x, train_y, alpha = 1)
plot(lasso_mod)
```

## Prepare training data in matrix form to be input to glmnet
```{r 3,include=TRUE}
train_x_matrix = model.matrix(~., data=train_x)
train_x_matrix=train_x_matrix[, colnames(train_x_matrix)!= "(Intercept)"] 
test_x_matrix = model.matrix(~., data=test_x)
test_x_matrix=test_x_matrix[, colnames(test_x_matrix)!= "(Intercept)"]
```

## Use 5-Fold CV when running LASSO using glmnet and default lambda grid
### Plot 5-Fold CV MSE against Lambda
```{r 4,include=TRUE,echo=TRUE}
set.seed(5)
cv5fold_lasso = cv.glmnet(train_x_matrix, train_y, alpha = 1,nfolds=5)
plot(cv5fold_lasso)
plot(cv5fold_lasso$lambda,cv5fold_lasso$cvm, 
     #  manually pull out lambdas and MSE 
     ## computed in this CV algorithm
     main="5-fold CV default settings - LASSO", 
     xlab="Lambda", ylab="CV MSE")
```

### Get Lambda with smallest CV MSE with default grid
```{r 5,include=TRUE}
minlambda_l = cv5fold_lasso$lambda[which.min(cv5fold_lasso$cvm)]
minlambda_l
```

## Get Test Set MSE with chosen lamda above
```{r 6,include=TRUE}
lasso_pred <- predict(lasso_mod, s = minlambda_l, newx = test_x_matrix)
mean((lasso_pred-test_y)^2)
```

## Use 5-Fold CV when running LASSO using glmnet and Self-defined lambda grid
## (We narrow down lambda)
## Plot 5-Fold CV MSE against Lambda
```{r 7,include=TRUE,echo=TRUE}
set.seed(5)
newgrid = seq(0, 5, length.out = 10000)
cv5fold_lasso_ug= cv.glmnet(train_x_matrix, train_y, alpha = 1, 
                             lambda=newgrid,nfolds=5)
plot(cv5fold_lasso_ug$lambda,cv5fold_lasso_ug$cvm, 
     #  manually pull out lambdas and MSE 
     ## computed in this CV algorithm
     main="5-fold CV - LASSO - User's Grid", 
     xlab="Lambda", ylab="CV MSE")
```

### Get Min Lambda with smallest CV MSE using  Self-defined lambda grid     
```{r 8,include=TRUE}
minlambda_l_ug = cv5fold_lasso_ug$lambda.min
minlambda_l_ug
```     
 
### Find the smallest 5-fold CV MSE for Min Lambda
```{r}
min(cv5fold_lasso_ug$cvm)
```

### Find the Standard Error of Lambda
```{r}
min(cv5fold_lasso_ug$cvsd)
```

## Gauge test set performance with fixed window using min lambda
```{r 9,include=TRUE}
lasso_pred <- predict(lasso_mod, s = minlambda_l_ug, newx = test_x_matrix)
mean((lasso_pred-test_y)^2)
```

### Get 1SE Lambda using  Self-defined lambda grid     
```{r ,include=TRUE}
minlambda_l_ug_1se = cv5fold_lasso_ug$lambda.1se
minlambda_l_ug_1se
```     
### Find the CV MSE for 1se Lambda
```{r}
cv5fold_lasso_ug$cvm[cv5fold_lasso_ug$lambda == minlambda_l_ug_1se]
```

## ## Gauge test set performance with fixed window using 1se lambda
```{r ,include=TRUE}
lasso_pred_lambda_1se <- predict(lasso_mod, s = minlambda_l_ug_1se, newx = test_x_matrix)
mean((lasso_pred_lambda_1se-test_y)^2)
```

# Get coefficient result for LASSO Min Lambda and LASSO 1SE Lambda
```{r}
coef = coef(cv5fold_lasso_ug, s=c(minlambda_l_ug,minlambda_l_ug_1se))
coef

```

# Expanding window forecasting

```{r}
df = combineddata
df_predict <- data.frame(Index=numeric(0),sg20y_true=numeric(0), forecastmin =numeric(0), 
  forecast1se = numeric(0), residmin = numeric(0), resid1se = numeric(0))

for (i in 144:177) {
  train.ew <- df[2:i,-1] # Creeping training size
  train.x.ew <-  model.matrix(sg20y ~ ., data = train.ew)[,-1] 
  train.y.ew <- train.ew$sg20y
  test.ew <- df[i+1,-1] 
  test.x.ew <- as.matrix(test.ew[,-1])
  true_sg20y <- test.ew$sg20y # Single test data
  
  set.seed(5)
  cv5fold.ew <- cv.glmnet(train.x.ew, train.y.ew, alpha=1, lambda = newgrid, nfolds = 5, type.measure = "mse")
  forecastmin <- predict(cv5fold.ew, newx=test.x.ew, s="lambda.min", type="response")
  forecast1se <- predict(cv5fold.ew, newx=test.x.ew, s="lambda.1se", type="response")
  
  df_predict[nrow(df_predict) + 1,] <- c(i+1,true_sg20y, forecastmin, forecast1se, 
    forecastmin - true_sg20y, forecast1se - true_sg20y)
} 


cat("Expanding window test MSE (lambda,min):", mean(df_predict$residmin^2),"\n",
  "Expanding window test MSE (lambda.1se):", mean(df_predict$resid1se^2))

```


## CV MSE of selected models using caret

```{r}
lassogrid = expand.grid("lambda" = cv5fold_lasso_ug$lambda.min, "alpha" = 1)
set.seed(5)
lambda_min_cv <- train(
  x = train_x, y = train_y, 
  method = "glmnet",
  trControl = trainControl(method = "cv", number = 5), tuneGrid = lassogrid
)
cat("CV MSE of LASSO (lambda.min) = ", mean(lambda_min_cv$resample$RMSE^2), "\n")

lassogrid1se = expand.grid("lambda" = cv5fold_lasso_ug$lambda.1se, "alpha" = 1)
set.seed(5)
lambda_1se_cv <- train(
  x = train_x, y = train_y, 
  method = "glmnet",
  trControl = trainControl(method = "cv", number = 5), tuneGrid = lassogrid1se
)
cat("CV MSE of LASSO (lambda.1se) = ", mean(lambda_1se_cv$resample$RMSE^2))

```
